---
title: "cm003"
author: "Wenzheng Zhou"
date: "October 31, 2017"
output: html_document
---

## Vector

```{r}
(v_log <- c(TRUE, FALSE, FALSE, TRUE))
#> [1]  TRUE FALSE FALSE  TRUE
(v_int <- 1:4)
#> [1] 1 2 3 4
(v_doub <- 1:4 * 1.2)
#> [1] 1.2 2.4 3.6 4.8
(v_char <- letters[1:4])
#> [1] "a" "b" "c" "d"
```

To "index a vector" means to address specific elements or atoms, either for reading or writing. We index a vector using square brackets, like so: x[something]. There are several ways to express which elements you want, i.e. there are several valid forms for something:To "index a vector" means to address specific elements or atoms, either for reading or writing. We index a vector using square brackets, like so: x[something]. There are several ways to express which elements you want, i.e. there are several valid forms for something:

- logical vector

```{r}
v_char[c(FALSE, FALSE, TRUE, TRUE)]
#> [1] "c" "d"
v_char[v_log]
#> [1] "a" "d"
```

- integer vector, all positive: the elements specified in something are kept

- negative integers, all negative: the elements specified in something are dropped

```{r}
#index by position
v_doub[2:3]
#> [1] 2.4 3.6
v_char[-4]
#> [1] "a" "b" "c"
```

- character vector: presumes that x is a named vector and the elements whose names are specified in something are kept not shown here, since none of our vectors are named

```{r}
v_int[0]
#> integer(0)
typeof(v_int[0])
#> [1] "integer"
v_doub[100]
#> [1] NA
typeof(v_doub[100])
#> [1] "double"
```

## list

```{r}
(x <- list(1:3, c("four", "five")))
#> [[1]]
#> [1] 1 2 3
#> 
#> [[2]]
#> [1] "four" "five"
(y <- list(logical = TRUE, integer = 4L, double = 4 * 1.2, character = "character"))
#> $logical
#> [1] TRUE
#> 
#> $integer
#> [1] 4
#> 
#> $double
#> [1] 4.8
#> 
#> $character
#> [1] "character"
(z <- list(letters[26:22], transcendental = c(pi, exp(1)), f = function(x) x^2))
#> [[1]]
#> [1] "z" "y" "x" "w" "v"
#> 
#> $transcendental
#> [1] 3.141593 2.718282
#> 
#> $f
#> function (x) 
#> x^2
```

Exercise: in list y, pull out the second and third elements in list z pull out all the transcendental numbers

```{r}
y[2:3]
z$transcendental
typeof(z["transcendental"])
z[["transcendental"]]
```

## vectorized operations

Consider how to square the integers 1 through n. The brute force method might look like:
```{r}
n <- 5
res <- rep(NA_integer_, n) 
for (i in seq_len(n)) {
  res[i] <- i ^ 2
}
res
```

The R way is 
```{r}
n <- 5
seq_len(n) ^ 2
#> [1]  1  4  9 16 25
```

Element-wise or vectorized operations are "baked in" to a surprising degree in R. Which is great. You get used to it soon enough.

But then there's the let down. This happens for atomic vectors, but not, in general, for lists. This makes sense because, in general, there is no reason to believe that the same sorts of operations make sense for each component of a list. Unlike atomic vectors, they are heterogeneous.

Here's a demo using as.list() to create the list version of an atomic vector.

```{r}
## elementwise exponentiation of numeric vector works
exp(v_doub)
#> [1]   3.320117  11.023176  36.598234 121.510418
## put the same numbers in a list and ... this no longer works :(
(l_doub <- as.list(v_doub))
#> [[1]]
#> [1] 1.2
#> 
#> [[2]]
#> [1] 2.4
#> 
#> [[3]]
#> [1] 3.6
#> 
#> [[4]]
#> [1] 4.8
exp(l_doub)
#> Error in exp(l_doub): non-numeric argument to mathematical function
```

So, how do you apply a function elementwise to a list?! What is the list analogue of exp(v_doub)?

Use purrr::map()! The first argument is the list to operate on. The second is the function to apply.

```{r}
library(purrr)
map(l_doub, exp)
#> [[1]]
#> [1] 3.320117
#> 
#> [[2]]
#> [1] 11.02318
#> 
#> [[3]]
#> [1] 36.59823
#> 
#> [[4]]
#> [1] 121.5104
```

use map and stringr fcn to split that have two words in them to two seprate character strings

```{r}
library(stringr)
fruit %>% map(str_split," ")
```

```{r}
fruit %>% map_chr(str_to_upper)
```

## Simplifying data from a list of GitHub users
```{r}
library(repurrrsive)
library(listviewer)
library(jsonlite)
library(dplyr)
library(tibble)
library(purrr)
```

The function .f will be [. And we finally get to use ...! This is where we pass the character vector of the names of our desired elements. We inspect the result for the first 2 users.

```{r}
map(gh_users, `[`, c("login", "name", "id", "location"))
```

